Concurrent Programming with CXL

### 2020-05-21 book/paper.tex
Latest
typos

So far we have pursued two properties: emph{mutual exclusion}
and \emph{progress}.  The former is an example of
\texttt{safety property}---it prevents something ``bad'' from
happening, like a reader and writer process both entering the
critical section.  The \emph{progress} property is an example
of a liveness property---that something good eventually happens.
Informally (and inexactly), progress states that if no processes
are in the critical section, then some process that wants to enter
can.

@@

While spinlocks suffer from starvation, it is a uniform random
process and each process has an equal chance of entering the critical
section.  Thus the probability of starvation is exponentially vanishing.
Unfortunately, such is not the case for the
reader/writer solution of Figure~\ref{fig:rw2lock}.
Consider this scenario: there are two readers and one writer.  One reader
is in the critical section while the writer is waiting.  Now the
second reader tries to enter and is able to.  The first reader leaves.
We are now in a similar situation as the initial state with one reader
in the critical section and the writer waiting, but it is not the same
reader.  Unfortunately for the writer, this scenario can repeat itself
indefinitely.  So even if neiter reader was in the critical section
for all of the time, and the second reader arrived well after the writer,
the writer never had a chance.
In this chapter, we will present a version of a reader/writer lock
implementation that solves this type of starvation, but won't eliminate
the type of starvation that comes with spinlocks.

>>>>

So far we have pursued two properties: \emph{mutual exclusion}
and \emph{progress}.  The former is an example of
\texttt{safety property}---it prevents something ``bad'' from
happening, like a reader and writer process both entering the
critical section.  The \emph{progress} property is an example
of a liveness property---that something good eventually happens.
Informally (and inexactly), progress states that if no processes
are in the critical section, then some process that wants to enter
can.

@@

While spinlocks suffer from starvation, it is a uniform random
process and each process has an equal chance of entering the critical
section.  Thus the probability of starvation is exponentially vanishing.
Unfortunately, such is not the case for the
reader/writer solution of Figure~\ref{fig:rw2lock}.
Consider this scenario: there are two readers and one writer.  One reader
is in the critical section while the writer is waiting.  Now the
second reader tries to enter and is able to.  The first reader leaves.
We are now in a similar situation as the initial state with one reader
in the critical section and the writer waiting, but it is not the same
reader.  Unfortunately for the writer, this scenario can repeat itself
indefinitely.  So even if neither reader was in the critical section
all of the time, and the second reader arrived well after the writer,
the writer never had a chance.
In this chapter, we will present a version of a reader/writer lock
implementation that solves this type of starvation, but won't eliminate
the type of starvation that comes with spinlocks.
